---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
library(tidyverse)
library(readr)

deaths <- av %>%
  pivot_longer(
    cols = starts_with("Death"),
    names_to = "Time",
    values_to = "Death"
  ) %>%
  mutate(Time = parse_number(Time)) %>%
  filter(Death != "") %>%
  select(-starts_with("Return"))

returns <- av %>%
  select(Name.Alias, starts_with("Return")) %>%
  pivot_longer(
    cols = starts_with("Return"),
    names_to = "Time",
    values_to = "Return"
  ) %>%
  mutate(Time = parse_number(Time)) %>%
  filter(Return != "")

av2 <- deaths %>%
  left_join(
    returns,
    by = c("Name.Alias", "Time"),
    relationship = "many-to-many"
  )
```

```{r}
av2 <- deaths %>%
  left_join(
    returns,
    by = c("Name.Alias", "Time"),
    relationship = "many-to-many"
  )
```

On average, an Avenger suffers 0.73 deaths. 


## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.


### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

## Nicole's Work:

## Scott's Work:"fans of the comics can expect one current or former member to die every seven months or so, with a permanent death occurring once every 20 months."

```{r}
# Count total deaths across Death1–Death5 columns
death_cols <- c("Death1", "Death2", "Death3", "Death4", "Death5")

total_deaths <- sum(av[death_cols] == "YES", na.rm = TRUE)

# Total time ~ 53 years
total_months <- 53 * 12

# Average months per death
months_per_death <- total_months / total_deaths

months_per_death


# Count permanent deaths
perm_deaths <- sum(av$Return1 == "NO", na.rm = TRUE)

# Total timeline
total_months <- 53 * 12

# Average time between permanent deaths
months_per_perm_death <- total_months / perm_deaths

months_per_perm_death
```
The average months between deaths is 7.15 which is in line with the claims of it being about 7 months. A permanent death is supposed to happen every 20 months on average but I am not getting that, I am getting 27.65 months which is longer than the stated average. If I had to guess it is something I am doing and not the original data analysis that is wrong.

## Chai's work: The data shows that Jocasta dies 5 times and returns 5 times. This matches the statement exactly, confirming that she has been destroyed five times and recovered five times

```{r}
jocasta_deaths <- deaths %>%
  filter(str_detect(Name.Alias, "Jocasta"), toupper(Death) == "YES") %>%
  nrow()

jocasta_returns <- returns %>%
  filter(str_detect(Name.Alias, "Jocasta"), toupper(Return) == "YES") %>%
  nrow()

jocasta_deaths
jocasta_returns
```



## Nicole's Work:


##Logan's work:
"Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team."

```{r}
length(unique(av2$URL))
```
There is 173 unique values for avengers so the number is right

```{r}
deathCounts <- av2 |> group_by(URL, Name.Alias) |> summarise(numDeaths = sum(Death == "YES"))
deathCounts
```
This is the number of deaths each avenger has

```{r}
atLeastOnce <- deathCounts |> filter(numDeaths > 0)
atLeastOnce
```
This gets all of the avengers that have died at least once (numDeaths != 0)

```{r}
dim(atLeastOnce)
```
This has 69 rows meaning that 69 people have died at least once since joining the Avengers. This means that the fact is correct.
